Setting up by calling relevant libraries and opening data file and assigning data variable:

```{r set up, results='hide'}
library(outliers)
library(ggplot2)
set.seed(93)
```

```{r openfile}
crimedata <- read.table('uscrime.txt',header = TRUE) #open data file with  headers then set header parameter to TRUE

head(crimedata) #First 6 rows of the data
```

Here, I am interested in the last column 'Crime' which accounts for number of crimes per 100,000 people.

```{r}
crime_number <- sort(crimedata$Crime, decreasing = TRUE) #I sorted this data in decreasing order

head(crime_number)
```

I want to find the outliers in the crime number data. Let's visualize that with a box and whisker plot.

```{r}
boxplot(crime_number, main="Crime Number per 100,000 People", col = "pink")
```

From the box and whisker plot, it seems that there are 3 outliers at the upper end of the data range. I will employ the Grubbs test to verify that.

However, before I begin the test, it should be noted that one of the assumptions of Grubbs test is normal distribution of data. Therefore, I will test for normality in the crime number data using Q-Q (quantile-quantile) plot. The Q-Q plot produces a 45 degree straight line representing how a normally distributed dataset would plot.

```{r}
qqnorm(crime_number)
qqline(crime_number)
```

Here, I observed that most data points of the crime number data are near to the Q-Q line with a few exceptions on the upper end. Therefore, I can assume the normality of the dataset to begin the Grubbs test.

The first Grubbs test will be testing the highest value of the range (i.e 1993)

```{r}
grubbs.test(crime_number)
```

The alternative hypothesis in the test is that 1993 is an outlier. Here, we have a p-value of 0.07887 which is arguably small enough if I consider a critical value of 0.08 in order for us to reject the null hypothesis in favor of the alternative. In this case, I can assume there is enough statistical evidence that 1993 is an outlier. 

I therefore removed 1993 from the crime_number data.

```{r}
crime_number <- crime_number[-1] #I was able to do this since I already sorted the data in decreasing order, the highest value has index of 1.
head(crime_number)
```

Next, I conduct the second Grubbs test for the second highest value in the original data.

```{r}
grubbs.test(crime_number)
```

In this test, the p-value is even lower at 0.02848, hence it is statistically significant enough to reject the null hypothesis and accep that alternative hypothesis that 1969 is an outlier.

As such, 1969 should also be removed from the data.

```{r}
crime_number <- crime_number[-1] #I was able to do this since I already sorted the data in decreasing order, the highest value has index of 1.
head(crime_number)
```

The box and whisker plot earlier showed that there are potentially 3 outliers. I haved detected 2, now I can run the test another to see if the last one is an outlier too.

```{r}
grubbs.test(crime_number)
```

Here, the p-value is 0.1781, large enough that I cannot reject the null hypothesis that 1674 is not an outlier. Therefore, I conclude that there is not enough statistical evidence that indicuates 1674 is an outlier.

Before I finished, I want to double check if the lower end of the data has any outlier even if the box and whisker plot suggested none.

```{r}
grubbs.test(crime_number, opposite = TRUE) #TRUE opposite checks for outliers on the other end of the data
```

I got a p-value of 1 and that is significant so I cannot reject the null hypothesis. Therefore, there is no outlier at the lower end of the crime number data. 

In conclusion, I found 2 outliers at the upper end of the crime number data and proceeded to remove then. After the removal, below is the crime number data that is ready for further analysis.

```{r}
crime_number
```

If the whole data frame is needed, here is the data frame that has the 2 outliers in crime number removed.

``` {r}
crimedata <- crimedata[order(crimedata$Crime, decreasing = TRUE),][-1:-2,] #removing the top 2 rows after sorting in decreasing order of crime number

head(crimedata)
```
